---
title: "Useful Packages"
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, message = FALSE, warning = FALSE)
options(tibble.print_min = 5, tibble.print_max = 5, max.print = 100L)
```

## R Packages

R comes bundled with so-called "base" packages, such as `tools::` and `utils::`. However, there are heaps of packages which are so widely used that they are considered _de facto_ base packages.

As standard, most of my scripts start as follows:

```{r}
library(tidyverse)
library(lubridate)
library(janitor)
library(readxl)
library(here)
```

If you want to install all of these packages, use the following code:

```{r, eval=FALSE}
install.packages(
  c(
    "tidyverse",
    "lubridate",
    "janitor",
    "readxl",
    "here",
  )
)

```

### `tidyverse`

The `tidyverse` is a so-called "meta-package", which actually installs a handful of useful packages such as `tidyr`, `readr`, and `dplyr`. 

To highlight one, `dplyr` is used for wrangling data frames (which the `tidyverse` calls "tibbles"). It can be used to filter rows, add columns, and aggregate data frames:

#### Filtering

```{r}
mtcars %>% 
  filter(cyl >= 8)
```

#### Adding Columns

```{r}
mtcars %>% 
  mutate(wt_kg = wt * 1000)
```

#### Summarising / Aggregating

```{r}
mtcars %>% 
  group_by(cyl) %>% 
  summarise(mean_hp = mean(hp))
```

The [`dplyr` cheat sheet is very useful](https://rstudio.com/wp-content/uploads/2015/02/data-wrangling-cheatsheet.pdf).

### `lubridate`

This is a package with nice functions for working with dates. Some notable ones are `day()`, `month()`, `year()`, which extract parts of the date; as well as `dmy()` (and similar), which coerce text into a date:

```{r}
today <- Sys.Date()

year(today)
month(today)
day(today)

dmy("20 Feb 2020")

```

### The Others

`janitor` provides some functions which tidy up data frames, `readxl` is for working with spreadsheets, and `here` is a simple way to get full file paths for local files.

If you want to learn more about them, either find them on GitHub and consult their README files, or run (for example) the following code to get a list of the vignettes associated with a package.

```{r, eval = FALSE}
vignette(package = "janitor")
```

Then you can open a specific vignette as follows:

```{r, eval = FALSE}
vignette("janitor")
```

### Learning

Probably the best place to learn more about the `tidyverse` and other useful packages is by reading the book [R For Data Science](https://r4ds.had.co.nz/), and by consulting the [cheatsheets](https://rstudio.com/resources/cheatsheets/) on RStudio's website.